---
title: "Querying a SQL database"
author: Pablo Barbera
---

## Creating a SQL database

For this first part of the materials, we'll be working with a SQLite database, which is self-contained in a file within our hard drive, without any need to set up a server. Note that even if our queries use this system, the language we use generalizes to a database of any size.

The data we will work with is a set of tweets by Members of the 114th U.S. Congress (2015-2016), which I collected from Twitter's public API..

We'll start by connecting to the database and adding the first table, with information at the member of congress level. 

```{r}
library(DBI)
```

```{r, eval=FALSE}
# create database: this will create a file in our hard drive where our tweets will be stored.
db <- dbConnect(RSQLite::SQLite(), "twitter-db.sqlite")

# reading the first file
congress <- read.csv("data/congress-twitter.csv",
	stringsAsFactors=F)

# adding first table: user-level data. 'name' indicates the name of the table
# where we will store our data, and 'value' is the data frame that we have
# loaded in memory with such table.
dbWriteTable(db, name="congress", value=congress)

# testing that it works with a simple query
dbListFields(db, "congress")
dbGetQuery(db, 'SELECT * FROM congress LIMIT 5')
```

Note that we have used `dbWriteTable` to add the data frame to our database, and then check that it all worked with `dbListFields` and `dbGetQuery`. As we will see in a second, this second function is what we will use to query the database, and it will always return a table.

We now switch to the second table. However, the files are too big to open them all in memory. Instead, we will open them one by one, and then __append__ them to the table. Let's see how that works...

```{r, eval=FALSE}
fls <- list.files("data/tweets", full.names=TRUE)

for (f in fls){
  
  message(f)

  # read file into memory
  tw <- read.csv(f, stringsAsFactors=F)
  
  # adding to table in SQL database
  dbWriteTable(db, "tweets", tw, append=TRUE)
  
}

# testing that it works
dbListFields(db, "tweets")
dbGetQuery(db, 'SELECT * FROM tweets LIMIT 5')

# what if we make a mistake and want to remove the table?
# dbRemoveTable(db, "tweets")

# and we close the connection for now
dbDisconnect(db)

```

## Querying an SQL database

Now that we have our tables in the database, let's see how we can query them. First we connect using `dbConnect` and then we type our query with `dbGetQuery`. Generally, we will write the query __clauses__ in capital letters and the column names in lowercase.

```{r}
db <- dbConnect(RSQLite::SQLite(), "twitter-db.sqlite")
test <- dbGetQuery(db, 'SELECT * FROM congress LIMIT 5')
str(test)
```

RMarkdown actually allows us to write the SQL query directly as a chunk of code. But it's much easier if we do this in R for now.

```{sql, connection=db}
SELECT * FROM congress LIMIT 5
```

This is the list of query clauses that we will use to work with the database table:

- __SELECT__: determines which columns to include in the query's result set

- __FROM__: identifies the tables from which to draw data and how tables should be joined

- __LIMIT__: in combination with SELECT, limits the number of records returned

- __WHERE__: filter out unwanted data

- __GROUP BY__: groups rows together by common column values

- __ORDER BY__: sorts the rows in the final result set by one or more columns


Let's start with some examples of __SELECT__:

```{r}
# querying just one column
dbGetQuery(db, "SELECT name FROM congress LIMIT 10")

# multiple columns
dbGetQuery(db, "SELECT name, party FROM congress LIMIT 10")

# adding expressions
dbGetQuery(db, "SELECT screen_name, favorite_count/retweet_count, UPPER(lang) FROM tweets LIMIT 10")

# adding aliases to the new columns
dbGetQuery(db, "SELECT screen_name, favorite_count/retweet_count AS like_ratio FROM tweets LIMIT 10")

# it's often a good idea to write a query in multiple lines so that it's easier to read it:
dbGetQuery(db, "SELECT screen_name, 
           favorite_count/retweet_count AS like_ratio 
           FROM tweets 
           LIMIT 10")

# selecting unique values from a column
dbGetQuery(db, "SELECT DISTINCT screen_name 
           FROM tweets 
           LIMIT 10")

```

Now some examples of how we can use __WHERE__ to subset the database based on values of specific columns:

```{r}
# selecting based on values of a column
dbGetQuery(db, "SELECT name, party 
           FROM congress
           WHERE party = 'Republican'
           LIMIT 10")

# greater than, working with dates
dbGetQuery(db, "SELECT screen_name, text, date 
           FROM tweets
           WHERE date > '2016-12-29'
           LIMIT 10")

# AND operator
dbGetQuery(db, "SELECT screen_name, text, date, retweet_count 
           FROM tweets
           WHERE date > '2016-11-01' AND retweet_count > 500
           LIMIT 10")

# OR operator
dbGetQuery(db, "SELECT screen_name, text, date, retweet_count, favorite_count
           FROM tweets
           WHERE date > '2016-12-01' AND (
              retweet_count > 100 OR favorite_count > 100
            )
           LIMIT 10")

# membership, IN
dbGetQuery(db, "SELECT screen_name, text, lang
           FROM tweets
           WHERE lang IN ('es', 'fr')
           LIMIT 10")

# matching conditions:
# _ matches exactly one character
# % matches any number of characters
dbGetQuery(db, "SELECT screen_name, text, date
           FROM tweets
           WHERE date LIKE '2015-01-__'
           LIMIT 10")

dbGetQuery(db, "SELECT screen_name, text, date
           FROM tweets
           WHERE date LIKE '2015-03%'
           LIMIT 10")

# SQLite does not have regular expressions, but we can be creative...
dbGetQuery(db, "SELECT screen_name, text, date
           FROM tweets
           WHERE text LIKE '%southern california%'
           LIMIT 1")

```


Very useful for us -- __GROUP BY__ allows us to aggregate results based on values of a specific variables. For example, what is the count of tweets by language?

```{r}
dbGetQuery(db, 
  "SELECT lang, COUNT(lang) AS lang_count
  FROM tweets
  GROUP BY lang")
```

And we can then order the results using __ORDER BY__. The default is alphabetical order for the first column, but we can change that:

```{r}
# sort by type_count
dbGetQuery(db, 
  "SELECT lang, COUNT(lang) AS lang_count
  FROM tweets
  GROUP BY lang
  ORDER BY lang_count")

# now in descending order
dbGetQuery(db, 
  "SELECT lang, COUNT(lang) AS lang_count
  FROM tweets
  GROUP BY lang
  ORDER BY lang_count DESC")

# which was the most popular tweet?
dbGetQuery(db, 
  "SELECT screen_name, text, date, retweet_count
  FROM tweets
  ORDER BY retweet_count DESC
  LIMIT 1")

# You can also specify the column number instead of the name
dbGetQuery(db, 
  "SELECT screen_name, text, date, retweet_count
  FROM tweets
  ORDER BY 4 DESC
  LIMIT 1")

# Who was the member of congress with most tweets?
dbGetQuery(db, 
  "SELECT screen_name, COUNT(screen_name) AS tweet_count
  FROM tweets
  GROUP BY screen_name
  ORDER BY tweet_count DESC
  LIMIT 5")

# and the least?
dbGetQuery(db, 
  "SELECT screen_name, COUNT(screen_name) AS tweet_count
  FROM tweets
  GROUP BY screen_name
  ORDER BY tweet_count
  LIMIT 5")

```






