---
title: "Querying a SQL database"
author: Pablo Barbera
---

## Querying multiple SQL tables

Now's where we get to see the magic of SQL. It makes it really easy to combine multiple SQL tables using __JOIN__. We just select columns from different tables and then specify what _key_ should be used to match them using __ON__. 

The example belows shows an __inner join__, but there are other types (left vs right outer joins, cross joins...) that we will not cover here.

```{r}
library(DBI)
db <- dbConnect(RSQLite::SQLite(), "twitter-db.sqlite")

dbGetQuery(db, 
  "SELECT tweets.retweet_count, congress.party, tweets.date, tweets.text
  FROM tweets JOIN congress
    ON congress.screen_name = tweets.screen_name
  LIMIT 10")
```

Note the structure: (1) to select the specific columns, you need to specify them as TABLENAME.COLUMNAME, (2) then set the tables you are merging using FROM FIRSTABLE JOIN SECONDTABLE, and finally (3) make sure you add which shared column should be used to merge: ON FIRSTABLE.KEY = SECONDTABLE.KEY.

Once we know how to merge tables, we can use columns from both tables. For example, let's say we want to find the most retweeted tweet by a Democrat, or by a Republican. Here's how we would do it:

```{r}
dbGetQuery(db, 
  "SELECT tweets.screen_name, tweets.text, tweets.retweet_count, congress.party 
  FROM tweets JOIN congress
    ON congress.screen_name = tweets.screen_name
  WHERE congress.party = 'Democrat'
  ORDER BY retweet_count DESC
  LIMIT 1")
dbGetQuery(db, 
  "SELECT tweets.screen_name, tweets.text, tweets.retweet_count, congress.party 
  FROM tweets JOIN congress
    ON congress.screen_name = tweets.screen_name
  WHERE congress.party = 'Republican'
  ORDER BY retweet_count DESC
  LIMIT 1")
```

Note that this will include retweets of retweets - how do we select only original tweets?

```{r}
dbGetQuery(db, 
  "SELECT tweets.screen_name, tweets.text, tweets.retweet_count, congress.party 
  FROM tweets JOIN congress
    ON congress.screen_name = tweets.screen_name
  WHERE congress.party = 'Democrat'
  AND tweets.text NOT LIKE 'RT @%'
  ORDER BY retweet_count DESC
  LIMIT 1")
dbGetQuery(db, 
  "SELECT tweets.screen_name, tweets.text, tweets.retweet_count, congress.party 
  FROM tweets JOIN congress
    ON congress.screen_name = tweets.screen_name
  WHERE congress.party = 'Republican'
  AND tweets.text NOT LIKE 'RT @%'
  ORDER BY retweet_count DESC
  LIMIT 1")
```


## Grouping and aggregating

SQL makes it really easy to compute aggregate statistics of one column based on values of a different column, with a series of built-in functions (COUNT, AVG, MAX, MIN...). Let's now see a few examples of this type of queries:

```{r}
# use COUNT with * to return the total number of rows
dbGetQuery(db, 'SELECT COUNT(*) FROM tweets')
dbGetQuery(db, 'SELECT COUNT(*) FROM congress')
```

Combining everything we've learned, we can for example count the total number of tweets by party, or by party and gender.

```{r}
dbGetQuery(db, 
  "SELECT congress.party, COUNT(*) AS total_tweets
  FROM tweets JOIN congress
    ON congress.screen_name = tweets.screen_name
  GROUP BY congress.party")

dbGetQuery(db, 
  "SELECT congress.party, congress.gender, COUNT(*) AS total_tweets
  FROM tweets JOIN congress
    ON congress.screen_name = tweets.screen_name
  GROUP BY congress.party, congress.gender")
```

There are other functions besides just __COUNT__. We can also compute averages, the maximum and minimu of a column, the sum... We can also functions like __SUBSTR__ to extract parts of a character string.

```{r}
# average number of tweets by party
dbGetQuery(db, 
  "SELECT congress.party, AVG(tweets.retweet_count), COUNT(*) 
  FROM tweets JOIN congress
    ON congress.screen_name = tweets.screen_name
  GROUP BY congress.party")

# we can also compute it in a different way...
dbGetQuery(db, 
  "SELECT congress.party, SUM(tweets.retweet_count)/COUNT(*) AS average 
  FROM tweets JOIN congress
    ON congress.screen_name = tweets.screen_name
  GROUP BY congress.party")

# most popular post by party
dbGetQuery(db, 
  "SELECT congress.screen_name, tweets.text, congress.party, MAX(tweets.retweet_count), COUNT(*) 
  FROM tweets JOIN congress
    ON congress.screen_name = tweets.screen_name
  GROUP BY congress.party")

# number of tweets by year-monnth
dbGetQuery(db, 
  "SELECT SUBSTR(date, 1, 7) AS year_month, COUNT(*) as post_count
  FROM tweets
  GROUP BY year_month")

```








