---
title: "Challenge 6: Machine learning"
author: Pablo Barbera
output: html_document
---

Now it's your turn to build a classifier! The file `incivility.csv` contains roughly a stratified random sample of 3,000 Facebook comments on the public pages of U.S. legislators. Each of these comments was labeled by a human annotator as containing `attacks` or not (`atacks` here is defined as: name-calling, insults, aspersions..., etc). We will take this as our working definition of incivility.

Read the dataset into R. What proportion of comments in this sample are classified as being uncivil?

```{r}
library(quanteda)
fb <- read.csv("data/incivility.csv", stringsAsFactors = FALSE)

```

Create a DFM for the comments on this dataset. Use sensible options (e.g. should you stem? trim it?)

```{r}

```

Now, split the dataset into training and test sets using similar code as what we just used.

```{r}


```

Build a LASSO classifier with the training set.

```{r}


```

Compute the accuracy, precision, and recall. Do you think this is a good classifier?

```{r}
## function to compute accuracy
accuracy <- function(ypred, y){
	tab <- table(ypred, y)
	return(sum(diag(tab))/sum(tab))
}
# function to compute precision
precision <- function(ypred, y){
	tab <- table(ypred, y)
	return((tab[2,2])/(tab[2,1]+tab[2,2]))
}
# function to compute recall
recall <- function(ypred, y){
	tab <- table(ypred, y)
	return(tab[2,2]/(tab[1,2]+tab[2,2]))
}


```

Find the features that are most predictive of each category. Do you think your classifier properly captured the concept of incivility?

```{r}


```

(Advanced) Do uncivil comments receive higher engagement? Use the `comments_likes_count` variable (measuring the number of likes on each comment) to answer this question. Use any statistical technique you may consider appropriate.

```{r}


```


