---
title: "GIS in R"
author: Pablo Barbera
date: October 31, 2017
output: html_document
---

### Working with shape files

The first part of this guided coding session will be devoted to learning how to work with shape files in R. The dataset we will contains the coordinates of around 16,000 tweets that mentioned the word "shutdown," collected on October 1st, 2013.

```{r}
df <- read.csv("../data/shutdown_locations.csv", stringsAsFactors=F)
str(df)
```

R already has some shape files built in that we can quickly work with. For example, at the country level:

```{r}
library(maps)
countries <- map.where(database="world", x=df$Longitude, y=df$Latitude)
tail(sort(table(countries)))
map("world")
```

At the state level in the U.S.:


```{r}
states <- map.where(database="state", x=df$Longitude, y=df$Latitude)
tail(sort(table(states)))
map("state")
```

And at the county level in the U.S:

```{r}
counties <- map.where(database="county", x=df$Longitude, y=df$Latitude)
tail(sort(table(counties)))
map("county")
```

But we can also load our own shape file. For example, the official region shape files, available [here](https://www.census.gov/geo/maps-data/data/tiger-cart-boundary.html).

```{r}
library(maptools)
region <- readShapePoly("../data/region/cb_2016_us_region_20m.shp")
plot(region)
region@data

# now convert to a Spatial Polygons object to identify region for each point
cc <- SpatialPolygons2map(region, namefield="NAME")
plot(cc) # this is what is internally in the object, just a set of points!
regions <- map.where(database=cc, x=df$Longitude, y=df$Latitude)
table(states)

```

### Producing choropleth map

Now imagine we want to produce a choropleth map, that is, a map where each area is shaded as a function of some parameter of the data. Let's say we want to produce a map of the U.S. states with each region shaded by the number of tweets.

First, we will create a data frame with the data

```{r}
states <- map.where(database="state", x=df$Longitude, y=df$Latitude)
tail(sort(table(states)))

# cleaning state names and creating data frame with counts
states <- gsub('(.*):.*', states, repl="\\1")
tab <- table(states, exclude=NULL)
tab <- data.frame(tab); names(tab) <- c("state", "tweets")
head(tab)
```

Now let's start with a simple map:

```{r}
library(ggplot2)
library(maps)
states_map <- map_data('state')

p <- ggplot(tab, aes(map_id=state))
p + geom_map(aes(fill=tweets), map = states_map) +
         expand_limits(x = states_map$long, y = states_map$lat)

```

We can then start playing around with the color and size of the borders.

```{r}
p + geom_map(aes(fill=tweets), map = states_map, color="yellow") +
         expand_limits(x = states_map$long, y = states_map$lat)

pq <- p + geom_map(aes(fill=tweets), map = states_map, color="grey", size=.1) +
         expand_limits(x = states_map$long, y = states_map$lat)
pq

```

And try with different color scales:

```{r}
pq + scale_fill_continuous(low="white", high="black") ## same as...

pq + scale_fill_gradient(low="white", high="black") ## smoothing btw 2 colors

pq + scale_fill_gradientn(colours = rainbow(7)) ## for >2 colors

pq + scale_fill_gradientn(colours = c("red", "white", "blue")) ## for >2 colors
 
library(RColorBrewer) ## by far the best. Check: http://colorbrewer2.org/
colors <- rev(brewer.pal(7,"YlOrRd"))

pq + scale_fill_continuous(low=colors[7], high=colors[1])
```

Given the skewed distribution of the dataset, we probably want to transform it by taking the log:

```{r}
pq + scale_fill_continuous(trans="log")

pq + scale_fill_continuous(trans="log", breaks=c(10, 100, 1000),
    labels = c("10", "100", "1K"))
```

The two final steps will be to remove axes and background, and to change the orientation and location of the legend

```{r}
pq <- pq + theme(
        # removing axes
        axis.line = element_blank(), axis.text = element_blank(), 
        axis.ticks = element_blank(), axis.title = element_blank(),
        # removing background / borders
        panel.background = element_blank(), plot.background = element_blank(),
        panel.border = element_blank(), panel.grid = element_blank()
    )
pq
```

```{r}
pq + theme(legend.direction="horizontal", 
    legend.position = c(.20, .10),
    legend.key.size = unit(c(.8), "cm"))

```


### Producing maps of locations

The other type of map we may want to produce is one where you overlay dots over a map. For example, let's produce a map of tweets in Europe, colored by language. The dataset we will use contains 80,300 tweets collected on September 15, 2013.

We will start by preparing the underlying map of Europe:

```{r}
library(cshapes)
library(scales)
library(ggplot2)
library(grid)
library(mapproj)

# preparing data frame with current country boundaries
world.data <- fortify(cshp(date=as.Date("2012-06-30")))

# base layer: map of Europe
p <- ggplot(world.data, aes(long,lat,group=group)) + 
    geom_polygon(fill="black", color="grey80", size=0.25) +
    coord_map("lagrange", xlim=c(-14.5, 35), ylim=c(32.5,61))
p

# removing axes, ticks, changing background...
pq <- p + theme(
    # dark background
        panel.background = element_rect(fill = "black"),
        plot.background = element_rect(fill="black"),
        # removing axis lines and ticks
        axis.line = element_blank(), axis.text = element_blank(), 
        axis.ticks = element_blank(), axis.title = element_blank(), 
        panel.border = element_blank(), 
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()
    )
pq
```

Now let's clean the tweet data:

```{r}
df <- read.csv("../data/loc_lang.csv", stringsAsFactors=F)
df <- df[df$lang!="und",] ## removing "undetermined" language
df <- df[df$lat>32.5 & df$lon>(-14.5) & df$lat<61 & df$lon<35,]
tab <- sort(table(df$lang), dec=TRUE) ## table for tweets by language
top15lang <- names(tab)[1:15] ## top 15 languages
df <- df[df$lang %in% top15lang,]
```

And we're ready to plot it!

```{r}
pq <- ggplot(world.data, aes(long,lat,group=group)) + 
    geom_polygon(fill="black", color="grey80", size=0.25) +
    geom_point(data = df, 
        aes(x = lon, y=lat, color=lang, group=NA), size=0.3) +
    coord_map("lagrange", xlim=c(-14.5, 35), ylim=c(32.5,61)) +
    theme(
        panel.background = element_rect(fill = "black"),
        plot.background = element_rect(fill="black"),
        axis.line = element_blank(), axis.text = element_blank(), 
        axis.ticks = element_blank(), axis.title = element_blank(), 
        panel.border = element_blank(), 
        panel.grid = element_blank(), 
        legend.background = element_rect(colour = F, fill = "black"),
        legend.key = element_rect(fill = "black", colour = F),
        legend.title = element_text(color="white"),
        legend.text = element_text(color="white", size=10)
    ) +
    guides(color = guide_legend(override.aes = list(size=5)))
# adding language labels
langs <- c("German", "English", "Spanish", "Estonian",
    "French", "Indonesian", "Italian", "Latvian", "Dutch",
    "Portuguese", "Russian", "Slovak", "Slovenian", 
    "Tagalog", "Turkish")
pq <- pq + scale_color_discrete("Language", labels = langs)
pq


```


